A Statistical Dynamical Model to Predict Extreme Events and Anomalous Features in Shallow Water Waves with Abrupt Depth Change

### 2018-11-25 paper1_tKdV/SI_appendix/SI_tKdV.tex
DQ

The TKdV equation [\ref{eq:tKdV}] can be also written in the spectral
domain for each Fourier mode $\hat{u}_{k}$ as
\[
\frac{d\hat{u}_{k}}{dt}=-\frac{ik}{2}E_{0}^{1/2}L_{0}^{-3/2}\sum_{\left|m\right|\leq\Lambda}\hat{u}_{m}\hat{u}_{m-k}^{*}+ik^{3}L_{0}^{-3}\hat{u}_{k}.
\]
It can be shown that the three conserved quantities above are still
conserved in this truncated system. With the spectral representation,
the conserved momentum and energy set the constraints on the spectral
coefficients 
\[
\mathcal{M}_{\Lambda}=\int_{0}^{2\pi}u_{\Lambda}dx=\left(2\pi\right)\hat{u}_{0}=0,\quad\mathcal{E}_{\Lambda}=\frac{1}{2}\int_{0}^{2\pi}\mathcal{P}_{\Lambda}\left(u_{\Lambda}^{2}\right)dx=2\pi\sum_{k=1}^{\Lambda}\left|\hat{u}_{k}\right|^{2}=1.
\]
The discretized Hamiltonian can be written accordingly as
\[
\mathcal{H}_{\Lambda}=E_{0}^{1/2}L_{0}^{-3/2}H_{3,\Lambda}-L_{0}^{-3}H_{2,\Lambda},\quad H_{3,\Lambda}=\frac{\pi}{3}\sum_{\left|m\right|,\left|n\right|\leq\Lambda}\hat{u}_{m}\hat{u}_{n}\hat{u}_{m+n}^{*},\;H_{2,\Lambda}=\pi\sum_{\left|k\right|\leq\Lambda}k^{2}\left|\hat{u}_{k}\right|^{2}.
\]
Especially, the Hamiltonian structure of the previous continuous equation
is maintained in this semi-discrete TKdV equation. The truncated equation
[\ref{eq:tKdV}] is still a Hamiltonian system with the corresponding
discrete Hamiltonian $\mathcal{H}_{\Lambda}$. Furthermore, the truncated
system [\ref{eq:tKdV}] satisfies the Liouville property \cite{abramov2003hamiltonian,majda2006nonlinear},
thus equilibrium statistical mechanism can be constructed based on
the conserved quantities.

@@

Here we summarize the direct numerical methods used for the validation
of the statistical theory in the \textcolor{blue}{\emph{main text}}. In general, the
strategy by using direct numerical simulations to confirm the anomalous
statistics follows the steps:
\begin{enumerate}
\item Generate samples from the incoming equilibrium distribution $\mu_{\infty}^{-}$
according to Algorithm \ref{alg:Metropolis-Hastings} for the incoming
waves. The model parameters $\left(E_{0},L_{0},\Lambda\right)$ are
determined from the experimental measurements and statistics in the
incoming flow;
\item Compute the expectation of the outflow Hamiltonian, $\left\langle H_{\Lambda}^{+}\right\rangle _{\mu_{\infty}^{-}}$,
under the chosen incoming Gibbs measure $\mu_{\infty}^{-}=\mathcal{G}_{\theta^{-}}$;
\item The outgoing flow statistics can be predicted by both the equilibrium
statistical mechanism and the direction model simulations. 
\begin{enumerate}
\item Equilibrium statistical mechanism: find the optimal outgoing flow
parameter $\theta_{*}^{+}$ from matching the expectation in [\ref{eq:matching}]
with the inflow value according to Algorithm \ref{alg:Recovering}.
Then the new equilibrium distribution $\mu_{\infty}^{+}$ after the
abrupt depth change can be sampled using the same method;
\item Direct model simulations: run the dynamical model [\ref{eq:KdV_simu}]
with outflow depth $D_{+}<1$ starting from the state sampled from
the inflow distribution $\mathcal{G}_{\theta^{-}}$. The outgoing
flow statistics can be captured by the solution from the direction
simulations. 
\end{enumerate}
\end{enumerate}
Above we proposed two approaches to predict the statistics in the
downstream wave disturbance. The equilibrium statistical mechanism
is an easy way to confirm the matching condition and get the equilibrium
distribution in far-field long-time limit. On the other hand, the
direct simulations of the dynamical model can offer estimates about
the transient statistics before the far-field equilibrium state is
reached. These two approaches can be used to confirm each other in
the numerical results.

@@

We compare the statistical spectra from the direct Monte Carlo model
simulation with the prediction from the equilibrium statistical mechanism
in Figure \ref{fig:Comparison-spec }. The flow energy spectra computed
both from the direct TKdV model simulations and from the Gibbs invariant
measure in the equilibrium statistical mechanism are compared. Starting
from two different $\theta^{-}$ initial states, the direct model
simulations offer the final equilibrium energy spectrum in a cimparable
decaying structure as predicted from the Gibbs measure and statistical
matching condition. This first comparison offers the consistency in
these two distinct approaches.

\begin{figure}
\begin{centering}
\includegraphics[scale=0.32]{./specEqui_th05}\includegraphics[scale=0.32]{./specEqui_th03}
\par\end{centering}
\caption{Comparison of the energy spectra computed from directly running the numerical 
model and the energy spectra predicted from the
Gibbs invariant measure in equilibrium statistical mechanism. The
model simulations start with the Gibbs measure with different
incoming flow parameter $\theta^{-}$.\label{fig:Comparison-spec }}
\end{figure}

>>>>

The TKdV equation [\ref{eq:tKdV}] can be also written in the spectral
domain for each Fourier mode $\hat{u}_{k}$ as
\[
\frac{d\hat{u}_{k}}{dt}=-\frac{ik}{2}E_{0}^{1/2}L_{0}^{-3/2}\sum_{\left|m\right|\leq\Lambda}\hat{u}_{m}\hat{u}_{m-k}^{*}+ik^{3}L_{0}^{-3}\hat{u}_{k}.
\]
It can be shown that the three conserved quantities above are still
conserved in this truncated system. With the spectral representation,
the conserved momentum and energy set the constraints on the spectral
coefficients 
\[
\mathcal{M}_{\Lambda}=\int_{0}^{2\pi}u_{\Lambda}dx=\left(2\pi\right)\hat{u}_{0}=0,\quad\mathcal{E}_{\Lambda}=\frac{1}{2}\int_{0}^{2\pi}\mathcal{P}_{\Lambda}\left(u_{\Lambda}^{2}\right)dx=2\pi\sum_{k=1}^{\Lambda}\left|\hat{u}_{k}\right|^{2}=1.
\]
The discretized Hamiltonian can be written accordingly as
\[
\mathcal{H}_{\Lambda}=E_{0}^{1/2}L_{0}^{-3/2}H_{3,\Lambda}-L_{0}^{-3}H_{2,\Lambda},\quad H_{3,\Lambda}=\frac{\pi}{3}\sum_{\left|m\right|,\left|n\right|\leq\Lambda}\hat{u}_{m}\hat{u}_{n}\hat{u}_{m+n}^{*},\;H_{2,\Lambda}=\pi\sum_{\left|k\right|\leq\Lambda}k^{2}\left|\hat{u}_{k}\right|^{2}.
\]
Especially, the Hamiltonian structure of the previous continuous equation
is maintained in this semi-discrete TKdV equation. The truncated equation
[\ref{eq:tKdV}] is still a Hamiltonian system with the corresponding
discrete Hamiltonian $\mathcal{H}_{\Lambda}$. Furthermore, the truncated
system [\ref{eq:tKdV}] satisfies the Liouville property \cite{abramov2003hamiltonian,majda2006nonlinear},
thus equilibrium statistical mechanics can be constructed based on
the conserved quantities.

@@

Here we summarize the direct numerical methods used for the validation
of the statistical theory in the \textcolor{blue}{\emph{main text}}. In general, the
strategy by using direct numerical simulations to confirm the anomalous
statistics follows the steps:
\begin{enumerate}
\item Generate samples from the incoming equilibrium distribution $\mu_{\infty}^{-}$
according to Algorithm \ref{alg:Metropolis-Hastings} for the incoming
waves. The model parameters $\left(E_{0},L_{0},\Lambda\right)$ are
determined from the experimental measurements and statistics in the
incoming flow;
\item Compute the expectation of the outflow Hamiltonian, $\left\langle H_{\Lambda}^{+}\right\rangle _{\mu_{\infty}^{-}}$,
under the chosen incoming Gibbs measure $\mu_{\infty}^{-}=\mathcal{G}_{\theta^{-}}$;
\item The outgoing flow statistics can be predicted by both the equilibrium
statistical mechanics and the direction model simulations. 
\begin{enumerate}
\item Equilibrium statistical mechanics: find the optimal outgoing flow
parameter $\theta_{*}^{+}$ from matching the expectation in [\ref{eq:matching}]
with the inflow value according to Algorithm \ref{alg:Recovering}.
Then the new equilibrium distribution $\mu_{\infty}^{+}$ after the
abrupt depth change can be sampled using the same method;
\item Direct model simulations: run the dynamical model [\ref{eq:KdV_simu}]
with outflow depth $D_{+}<1$ starting from the state sampled from
the inflow distribution $\mathcal{G}_{\theta^{-}}$. The outgoing
flow statistics can be captured by the solution from the direction
simulations. 
\end{enumerate}
\end{enumerate}
Above we proposed two approaches to predict the statistics in the
downstream wave disturbance. The equilibrium statistical mechanics
is an easy way to confirm the matching condition and get the equilibrium
distribution in far-field long-time limit. On the other hand, the
direct simulations of the dynamical model can offer estimates about
the transient statistics before the far-field equilibrium state is
reached. These two approaches can be used to confirm each other in
the numerical results.

@@

We compare the statistical spectra from the direct Monte Carlo model
simulation with the prediction from the equilibrium statistical mechanics
in Figure \ref{fig:Comparison-spec }. The flow energy spectra computed
both from the direct TKdV model simulations and from the Gibbs invariant
measure in the equilibrium statistical mechanics are compared. Starting
from two different $\theta^{-}$ initial states, the direct model
simulations offer the final equilibrium energy spectrum in a cimparable
decaying structure as predicted from the Gibbs measure and statistical
matching condition. This first comparison offers the consistency in
these two distinct approaches.

\begin{figure}
\begin{centering}
\includegraphics[scale=0.32]{./specEqui_th05}\includegraphics[scale=0.32]{./specEqui_th03}
\par\end{centering}
\caption{Comparison of the energy spectra computed from directly running the numerical 
model and the energy spectra predicted from the
Gibbs invariant measure in equilibrium statistical mechanics. The
model simulations start with the Gibbs measure with different
incoming flow parameter $\theta^{-}$.\label{fig:Comparison-spec }}
\end{figure}

### 2018-11-25 paper1_tKdV/tKdV_theory.tex
DQ

Recent laboratory experiments in weakly turbulent shallow water reveal a remarkable transition from Gaussian to anomalous behavior as surface waves cross an abrupt depth change (ADC). 
A normally-distributed incoming wave train, upstream of the ADC, transitions to a highly non-Gaussian outgoing wave train, downstream of the ADC, that exhibits large positive skewness of the surface height and more frequent extreme events \cite{bolles2018anomalous}.
Here we develop a statistical dynamical model to explain and quantitatively
predict this anomalous behavior as experimental control
parameters are varied. The first step is to use incoming and outgoing
truncated Korteweg-de Vries (TKdV) equations matched in time at the
ADC. The TKdV equation is a Hamiltonian system which induces incoming
and outgoing statistical Gibbs invariant measures which are statistically
matched at the ADC. The statistical matching of the known nearly Gaussian
incoming Gibbs state at the ADC completely determines the predicted
anomalous outgoing Gibbs state, which can be calculated by a simple
MCMC algorithm, verified by direct numerical simulations, and successfully
captures key features of the experiment. There is even an analytic
formula for the anomalous outgoing skewness. The strategy here should
be useful for predicting extreme anomalous statistical behavior in
other dispersive media in different settings.

\todo[inline]{If you like, we could cite papers on `Optical rogue waves' and/or rogue waves in microwave systems here.}

@@

Controlled laboratory experiments were carried out in  \cite{bolles2018anomalous} to examine the statistical behavior of surface waves crossing an ADC. In these experiments, nearly unidirectional waves are generated by a paddle wheel and propagate through a long, narrow wave tank. Midway through, the waves encounter a step in the bottom topography, and thus abruptly transition from one depth to another. The paddle wheel is forced with angle
\[
\theta\left(t\right)=\theta_{0}+\Delta\theta\sum_{n=1}^{N}a_{n}\cos\left(\omega_{n}t+\delta_{n}\right),\;E\sim\left(\Delta\theta\right)^{2}\sum_{n}a_{n}^{2}\omega_{n}^{2}.
\]
where the weights $a_n$ are Gaussian in spectral space. The peak frequency $\omega_p$ gives rise to a characteristic wavelength $\lambda_c$ of water waves, which can be estimated using the dispersion relation.
Above, the total energy $E$ in the system is determined by the angle amplitude
$\Delta\theta$. Optical measurements of the free surface reveal a number of surprising statistical features:
\begin{itemize}
\item Distinct statistics are found between the incoming and outgoing wave
disturbances: incoming waves display near-Gaussian statistics,
while outgoing waves skew strongly towards positive displacement.
\item The degree of non-Gaussianity present in the outgoing wave field depends on the total energy of the system: 
larger driving amplitudes $\Delta\theta$ generate stronger skewness in the surface displacement PDFs.
\item The waves also show different characteristic peak wavelengths in incoming
and outgoing flows.

\todo[inline]{I am not sure we need to include this last item. It is not really an experimental observation, simply a calculation that can be done using the dispersion relation and the values of the two depths. Could we simply take this last item out? Would you prefer to replace it with another observation? I could easily come up with one. Perhaps a comment on the decay rate of the power spectra.}

@@

\todo[inline]{I modified the last sentence above. Feel free to revert to your original sentence if you do not like mine.}

@@

The model parameters $\left(E_{0},L_{0},\Lambda\right)$ in [\ref{eq:tKdV_nondim}]
can be directly linked to the basic scales from the physical problem.
The important characterizing parameters measured from the experiments
include: $\epsilon=\frac{a}{H_{0}}$ the wave amplitude $a$ to water
depth $H_{0}$ ratio; $\delta=\frac{H_{0}}{\lambda_{c}}$ the water
depth to wavelength scale $\lambda_{c}$ ratio; and $D_{0}=\frac{d}{H_{0}}$
the normalized wave depth ratio with incoming flow depth $d=H_{0}$
to the outgoing flow depth $d<H_{0}$. The interpretations and reference
values of these model parameters are based on the experimental setup
\cite{bolles2018anomalous}. By comparing the characteristic physical
scales, the normalized TKdV equation [\ref{eq:tKdV_nondim}] can be
linked directly with the measured non-dimensional quantities by
\begin{equation}
L_{0}=6^{\frac{1}{3}}\left(M\epsilon^{\frac{1}{2}}\delta^{-1}\right),\;E_{0}=\frac{27}{2}\gamma^{-2}\left(M\epsilon^{\frac{1}{2}}\delta^{-1}\right),\label{eq:params}
\end{equation}
where $M$ defines the computational domain size $L_{d}=M\lambda_{c}$
as $M$-multiple of the characteristic wavelength $\lambda_{c}$,
and $\gamma=\frac{U}{a}$ represents the factor to normalize the total
energy in the state variable $u_{\Lambda}$ to one.

\todo[inline]{Is there a difference between $L_0$ and $L_d$? typo?}

@@

\section{Equilibrium statistical mechanism for generating the stationary invariant measure}

\todo[inline]{In this section title, did you mean `statistical mechanism' or `statistical mechanics'?}

@@

The Gibbs measures $\mathcal{G}_{\theta}^{\pm}$ are defined based
on the different inverse temperatures $\theta^{\pm}$ on the two sides
of the solutions
\[
\begin{aligned}\mu_{t}^{-}\left(u_{\Lambda}^{-};D_{-}\right), & \quad u_{\Lambda}\mid_{t=T_{\mathrm{ADC}-}}=u_{0},\;t<T_{\mathrm{ADC}};\\
\mu_{t}^{+}\left(u_{\Lambda}^{+};D_{+}\right), & \quad u_{\Lambda}\mid_{t=T_{\mathrm{ADC}+}}=u_{0},\;t>T_{\mathrm{ADC}},
\end{aligned}
\]
where $u_{0}$ represents the deterministic matching condition between
the incoming and outgoing waves. The two distributions, $\mu_{t}^{-},\mu_{t}^{+}$
should also be matched at the depth change location $T_{\mathrm{ADC}}$,
so that
\[
\mu_{\infty}^{-}\left(u_{\Lambda}\right)=\mu_{t=T_{\mathrm{ADC}}}^{-}\left(u_{\Lambda}\right)=\mu_{t=T_{\mathrm{ADC}}}^{+}\left(u_{\Lambda}\right).
\]
\todo[inline]{I do not believe we have defined $\mu_{\infty}^{-}$ anywhere.}
In matching the flow statistics before and after the abrupt depth
change, we first use the conservation of the deterministic Hamiltonian
$H_{\Lambda}^{+}$ after the depth change. Then assuming ergodicity
\cite{abramov2003hamiltonian,majda2006nonlinear}, the statistical
expectation for the Hamiltonian $\left\langle H_{\Lambda}^{+}\right\rangle $
is conserved in time after the depth change at $t=T_{\mathrm{ADC}}$
and should stay in the same value as the system approaches equilibrium
as $t\rightarrow\infty$. The final statistical matching condition
to get the outgoing flow statistics with parameter $\theta^{+}$ can
be found by
\begin{equation}
\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta}^{+}}=\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta}^{-}},\label{eq:matching}
\end{equation}
with the outgoing flow Hamiltonian $H_{\Lambda}^{+}$ and the Gibbs
measures $\mathcal{G}_{\theta}^{\pm}$ before and after the abrupt
depth change.

@@

As a result, the downstream statistics in final equilibrium predicted
from the direct numerical simulations here agree with the equilibrium
statistical mechanism prediction illustrated in Figure \ref{fig:matching}. The prediction from these two different approaches
confirm each other.
\todo[inline]{Again, do we mean statistical `mechanism' or `mechanics'?}

@@

\begin{figure*}[h]
\centering
\subfloat[downstream state $D_{0}=0.24$]{\includegraphics[scale=.95]{./fig3_1}}
\vspace{-1em}
\subfloat[upstream state $D_{0}=1$]{\includegraphics[scale=.95]{./fig3_2}}
\vspace{-1em}
\caption{Realization of the downstream and upstream flow solutions $u_{\Lambda}^{\pm}$. Note the larger vertical scale in the downstream time-series plot.\label{fig:Realization}}
\end{figure*}
\begin{itemize}
\item The transition from near-Gaussian to skewed non-Gaussian distribution
as well as the jump in both skewness and kurtosis observed in the
experiment observations (Fig. 1 of  \cite{bolles2018anomalous}) can
be characterized by the statistical model simulation results (see
the first and second row of Figure \ref{fig:Changes-stats}). Notice
that the difference in the decay of third and fourth moments in the far
end of the downstream regime from the experimental data is due to the
dissipation effect in the flow from the wave absorbers that is not modeled in the statistical
model here. The model simulation time-series plotted in Figure \ref{fig:Realization}
can be compared with the observed time sequences from experiments
(Fig. 1 of \cite{bolles2018anomalous}). The downstream simulation generates
waves with strong and frequent intermittency towards the positive
displacement, while the upstream waves show symmetric displacements
in two directions with at most small peaks in slow time. Even in the
time-series at a single location $x=0$, the long-time variation displays similar
structures. 
\item The downstream PDFs in experimental data are estimated with a Gamma
distribution in Fig. 2 of \cite{bolles2018anomalous}. Here in the same
way, we can fit the highly skewed outgoing flow PDFs from the numerical
results with the Gamma distribution
\[
\rho\left(u;k,\alpha\right)=\frac{e^{-k}\alpha^{-1}}{\Gamma\left(k\right)}\left(k+\alpha^{-1}u\right)^{k-1}e^{-\alpha^{-1}u}.
\]
The parameters $\left(k,\alpha\right)$ in the Gamma distribution
are fitted according to the measured statistics in skewness and variance,
that is, $\sigma^{2}=k\alpha^{2},\kappa_{3}=2/\sqrt{k}$. And the excess
kurtosis of the Gamma distribution can be recovered as $\kappa_{4}=6/k$.
As shown in the third row of Figure \ref{fig:Changes-stats}, excellent
agreement in the PDFs with the Gamma distributions
is reached in consistency with the experimental data observations.
The accuracy with this approximation increases as the initial inverse
temperature $\theta^{-}$ increases in value to generate more skewed
distribution functions.
\item The experiments also have the up and down stream power spectra in
time (Fig. 4 of \cite{bolles2018anomalous}), which shows more energy
at small time scales, i.e., a relatively slower decay rate in the
downstream compared with the upstream case. This is also observed
in the direct numerical simulations here (detailed results shown
in \textcolor{blue}{\emph{SI Appendix, C.2}}). The downstream state contains
more energetic high frequencies.
The peak frequency illustrates the occurrence of the transporting
waves along the water tank.

>>>>

Recent laboratory experiments in weakly turbulent shallow water reveal a remarkable transition from Gaussian to anomalous behavior as surface waves cross an abrupt depth change (ADC). 
A normally-distributed incoming wave train, upstream of the ADC, transitions to a highly non-Gaussian outgoing wave train, downstream of the ADC, that exhibits large positive skewness of the surface height and more frequent extreme events \cite{bolles2018anomalous}.
Here we develop a statistical dynamical model to explain and quantitatively
predict this anomalous behavior as experimental control
parameters are varied. The first step is to use incoming and outgoing
truncated Korteweg-de Vries (TKdV) equations matched in time at the
ADC. The TKdV equation is a Hamiltonian system which induces incoming
and outgoing statistical Gibbs invariant measures which are statistically
matched at the ADC. The statistical matching of the known nearly Gaussian
incoming Gibbs state at the ADC completely determines the predicted
anomalous outgoing Gibbs state, which can be calculated by a simple
MCMC algorithm, verified by direct numerical simulations, and successfully
captures key features of the experiment. There is even an analytic
formula for the anomalous outgoing skewness. The strategy here should
be useful for predicting extreme anomalous statistical behavior in
other dispersive media in different settings  \cite{solli2007optical,hohmann2010freak}.

\todo[inline]{If you like, we could cite papers on `Optical rogue waves' and/or rogue waves in microwave systems here. (references added)}

@@

Controlled laboratory experiments were carried out in  \cite{bolles2018anomalous} to examine the statistical behavior of surface waves crossing an ADC. In these experiments, nearly unidirectional waves are generated by a paddle wheel and propagate through a long, narrow wave tank. Midway through, the waves encounter a step in the bottom topography, and thus abruptly transition from one depth to another. The paddle wheel is forced with angle
\[
\theta\left(t\right)=\theta_{0}+\Delta\theta\sum_{n=1}^{N}a_{n}\cos\left(\omega_{n}t+\delta_{n}\right),\;E\sim\left(\Delta\theta\right)^{2}\sum_{n}a_{n}^{2}\omega_{n}^{2}.
\]
where the weights $a_n$ are Gaussian in spectral space. The peak frequency $\omega_p$ gives rise to a characteristic wavelength $\lambda_c$ of water waves, which can be estimated using the dispersion relation.
Above, the total energy $E$ in the system is determined by the angle amplitude
$\Delta\theta$. Optical measurements of the free surface reveal a number of surprising statistical features:
\begin{itemize}
\item Distinct statistics are found between the incoming and outgoing wave
disturbances: incoming waves display near-Gaussian statistics,
while outgoing waves skew strongly towards positive displacement.
\item The degree of non-Gaussianity present in the outgoing wave field depends on the total energy of the system: 
larger driving amplitudes $\Delta\theta$ generate stronger skewness in the surface displacement PDFs.
\item \textcolor{blue}{The waves also show different slopes in the power spectra at upstream and downstream locations}.

\todo[inline]{I am not sure we need to include this last item. It is not really an experimental observation, simply a calculation that can be done using the dispersion relation and the values of the two depths. Could we simply take this last item out? Would you prefer to replace it with another observation? I could easily come up with one. Perhaps a comment on the decay rate of the power spectra. (I will confirm with Andy whether he would just delete this item or keep something else like the one above)}

@@

\todo[inline]{I modified the last sentence above. Feel free to revert to your original sentence if you do not like mine. (the change looks good)}

@@

The model parameters $\left(E_{0},L_{0},\Lambda\right)$ in [\ref{eq:tKdV_nondim}]
can be directly linked to the basic scales from the physical problem.
The important characterizing parameters measured from the experiments
include: $\epsilon=\frac{a}{H_{0}}$ the wave amplitude $a$ to water
depth $H_{0}$ ratio; $\delta=\frac{H_{0}}{\lambda_{c}}$ the water
depth to wavelength scale $\lambda_{c}$ ratio; and $D_{0}=\frac{d}{H_{0}}$
the normalized wave depth ratio with incoming flow depth $d=H_{0}$
to the outgoing flow depth $d<H_{0}$. The interpretations and reference
values of these model parameters are based on the experimental setup
\cite{bolles2018anomalous}. By comparing the characteristic physical
scales, the normalized TKdV equation [\ref{eq:tKdV_nondim}] can be
linked directly with the measured non-dimensional quantities by
\begin{equation}
L_{0}=6^{\frac{1}{3}}\left(M\epsilon^{\frac{1}{2}}\delta^{-1}\right),\;E_{0}=\frac{27}{2}\gamma^{-2}\left(M\epsilon^{\frac{1}{2}}\delta^{-1}\right),\label{eq:params}
\end{equation}
where $M$ defines the computational domain size $M\lambda_{c}$
as $M$-multiple of the characteristic wavelength $\lambda_{c}$,
and $\gamma=\frac{U}{a}$ represents the factor to normalize the total
energy in the state variable $u_{\Lambda}$ to one.

\todo[inline]{Is there a difference between $L_0$ and $L_d$? typo? (These two are actually different. $L_d$ is the `true' length scale defined in SI eqn. [10]. I guess here it's better to delete this notation to avoid confusion.)}

@@

\section{Equilibrium statistical \textcolor{blue}{mechanics} for generating the stationary invariant measure}

\todo[inline]{In this section title, did you mean `statistical mechanism' or `statistical mechanics'? (typos corrected)}

@@

The Gibbs measures $\mathcal{G}_{\theta}^{\pm}$ are defined based
on the different inverse temperatures $\theta^{\pm}$ on the two sides
of the solutions
\[
\begin{aligned}\mu_{t}^{-}\left(u_{\Lambda}^{-};D_{-}\right), & \quad u_{\Lambda}\mid_{t=T_{\mathrm{ADC}-}}=u_{0},\;t<T_{\mathrm{ADC}};\\
\mu_{t}^{+}\left(u_{\Lambda}^{+};D_{+}\right), & \quad u_{\Lambda}\mid_{t=T_{\mathrm{ADC}+}}=u_{0},\;t>T_{\mathrm{ADC}},
\end{aligned}
\]
where $u_{0}$ represents the deterministic matching condition between
the incoming and outgoing waves. The two distributions, $\mu_{t}^{-},\mu_{t}^{+}$
should also be matched at the depth change location $T_{\mathrm{ADC}}$,
so that,
\[
\mu_{t=T_{\mathrm{ADC}}}^{-}\left(u_{\Lambda}\right)=\mu_{t=T_{\mathrm{ADC}}}^{+}\left(u_{\Lambda}\right).
\]
\todo[inline]{I do not believe we have defined $\mu_{\infty}^{-}$ anywhere. (removed to avoid confusion)}
In matching the flow statistics before and after the abrupt depth
change, we first use the conservation of the deterministic Hamiltonian
$H_{\Lambda}^{+}$ after the depth change. Then assuming ergodicity
\cite{abramov2003hamiltonian,majda2006nonlinear}, the statistical
expectation for the Hamiltonian $\left\langle H_{\Lambda}^{+}\right\rangle $
is conserved in time after the depth change at $t=T_{\mathrm{ADC}}$
and should stay in the same value as the system approaches equilibrium
as $t\rightarrow\infty$. The final statistical matching condition
to get the outgoing flow statistics with parameter $\theta^{+}$ can
be found by
\begin{equation}
\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta}^{+}}=\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta}^{-}},\label{eq:matching}
\end{equation}
with the outgoing flow Hamiltonian $H_{\Lambda}^{+}$ and the Gibbs
measures $\mathcal{G}_{\theta}^{\pm}$ before and after the abrupt
depth change.

@@

As a result, the downstream statistics in final equilibrium predicted
from the direct numerical simulations here agree with the equilibrium
statistical mechanics prediction illustrated in Figure \ref{fig:matching}. The prediction from these two different approaches
confirm each other.
\todo[inline]{Again, do we mean statistical `mechanism' or `mechanics'? (should be `mechanics')}

@@

\begin{figure*}[h]
\centering
\subfloat[downstream state $D_{0}=0.24$]{\includegraphics[scale=.95]{./fig3_1}}
\vspace{-1em}
\subfloat[upstream state $D_{0}=1$]{\includegraphics[scale=.95]{./fig3_2}}
\vspace{-1em}
\caption{Realization of the downstream and upstream flow solutions $u_{\Lambda}^{\pm}$. Note the larger vertical scale in the downstream time-series plot.\label{fig:Realization}}
\end{figure*}
\begin{itemize}
\item The transition from near-Gaussian to skewed non-Gaussian distribution
as well as the jump in both skewness and kurtosis observed in the
experiment observations (Fig. 1 of  \cite{bolles2018anomalous}) can
be characterized by the statistical model simulation results (see
the first and second row of Figure \ref{fig:Changes-stats}). Notice
that the difference in the decay of third and fourth moments in the far
end of the downstream regime from the experimental data is due to the
dissipation effect in the flow from the wave absorbers that is not modeled in the statistical
model here. The model simulation time-series plotted in Figure \ref{fig:Realization}
can be compared with the observed time sequences from experiments
(Fig. 1 of \cite{bolles2018anomalous}). The downstream simulation generates
waves with strong and frequent intermittency towards the positive
displacement, while the upstream waves show symmetric displacements
in two directions with at most small peaks in slow time. Even in the
time-series at a single location $x=0$, the long-time variation displays similar
structures. 
\item The downstream PDFs in experimental data are estimated with a Gamma
distribution in Fig. 2 of \cite{bolles2018anomalous}. Here in the same
way, we can fit the highly skewed outgoing flow PDFs from the numerical
results with the Gamma distribution
\[
\rho\left(u;k,\alpha\right)=\frac{e^{-k}\alpha^{-1}}{\Gamma\left(k\right)}\left(k+\alpha^{-1}u\right)^{k-1}e^{-\alpha^{-1}u}.
\]
The parameters $\left(k,\alpha\right)$ in the Gamma distribution
are fitted according to the measured statistics in skewness and variance,
that is, $\sigma^{2}=k\alpha^{2},\kappa_{3}=2/\sqrt{k}$. And the excess
kurtosis of the Gamma distribution can be recovered as $\kappa_{4}=6/k$.
As shown in the third row of Figure \ref{fig:Changes-stats}, excellent
agreement in the PDFs with the Gamma distributions
is reached in consistency with the experimental data observations.
The accuracy with this approximation increases as the initial inverse
temperature $\theta^{-}$ increases in value to generate more skewed
distribution functions.
\item The experiments also have the up and down stream power spectra in
time (Fig. 4 of \cite{bolles2018anomalous}), which shows more energy
at small time scales, i.e., a relatively slower decay rate in the
downstream compared with the upstream case. This is also observed
in the direct numerical simulations here (detailed results shown
in \textcolor{blue}{\emph{SI Appendix, C.2}}). The downstream state contains
more energetic high frequencies.
\textcolor{blue}{The peak frequency illustrates the characteristic time scale of the transporting
wave trains along the water tank}.
